---
title: "MTH522 Homework 1"
author: "Anubhav Shankar"
date: "2022-10-03"
output:
  html_document: default
  pdf_document: default
---




```{r Install the relevant packages, results='hide'}
library(UsingR)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(rmarkdown)
library(knitr)
```





We'll now load the **Pearson's Father-Son Height** data for further evaluation



```{r Load data}
fason <- father.son # Load the data and save it into an object
glimpse(fason) # Get a quick overview of the data
str(fason) # Same as glimpse()
head(fason) # Preliminary exploration

```





So, the data contains of two numerical features - **Father's Height (fheight)** and **Son's Height (sheight)**. There are a total of 1078 entries.





Let's create a simple linear regression model by making the son's height the dependent variable and father's height as the independent variable.



```{r Linear Regression}
model_fit <- lm(fason$sheight ~ fason$fheight) # Create a simple Linear Regression Model
summary(model_fit) # Get the model statistics
```




**The Goodness of Fit (R^2) for this data is only 0.2506 i.e. only 25% of the data variance is explained by the independent variable, father's height in this case. Also, we see that father's height is a significant variable having a p-value << 0.05**




Let us now create a simple scatter plot to see the relationship between the two variables.

```{r Simple Scatter Plot,results='hide'}
plot(fason$fheight, fason$sheight, xlab = "Father's Height(in.)", ylab = "Son's Height(in.)", 
     pch = 20) + title("Height Comparison") # Create a simple Scatter Plot 
```





We see that there is a strong concentration of the observations. This might indicate some *correlation*. However, let's create a more detailed plot by adding a regression line and a SD line along with plotting the respective means of the heights. 




```{r A more detailed plot using ggplot2}
plot.new() # Use this when creating a .Rmd to prevent errors if you have multiple plots in a single file
ggplot(fason,aes(x = fheight, y = sheight)) + # This line initiates the plotting function
geom_smooth(method = "lm", color = "red", size = 2) + # This line creates the regression line
geom_point() + # This line generates the points of the scatter plot
points(mean(fason$fheight),mean(fason$sheight)) + # This line basically sets the coordinates
geom_vline(xintercept = mean(fason$fheight),color = "green") + # Adds a vertical line passing through the mean of father's height
geom_hline(yintercept = mean(fason$sheight),color = "green") + # Adds a vertical line passing through the mean of son's height
geom_abline(slope = model_fit$coefficients[2],intercept = model_fit$coefficients[1],color = "blue",size = 1) + # This line creates the SD line by getting the intercept and slope from the regression model
xlab("Father's Height") + # Adds a x-axis label
ylab("Son's Height") # Adds a y-axis label
```



We see that the **SD line and the regression line overlap perfectly**. Concurrently, the mean of father's height and the son's height are pretty close by as well.